Extended Data

### 2022-08-02 doc/manuscript/Dow_manuscript_Nature.Rmd
fix typo in equation
removed extra closing parentheses

We fit a five-parameter logistic growth model[@mcmahon_general_2015] to dendrometer band data from each tree-year to define stem growth milestones and growth rates (Fig. `r schematic_summary`). 
In particular, we model the observed diameter at breast height (DBH) on a given day of the year (DOY; *i.e.*, julian days) as: 
$$ DBH = L + \frac{K - L} { 1 + 1/\theta \cdot exp[-r(DOY - DOY_{ip}) / \theta)] ^ \theta}$$ 
Here, _L_ and _K_ are lower and upper asymptotes of the model. 
*DOY~ip~* is the day of year where the inflection point in growth rate occurs,
_r_ shapes the slope of the curve at the inflection point,
and $\theta$ is a tuning parameter controlling the slope of the curve toward the upper asymptote.
This allows an asymmetric fit to the data, where the onset of growth can be estimated independent of the cessation of growth. 
When $\theta = 1$, *g~max~* occurs on *DOY~ip~*.
The model outputs two additional variables, _a_ and _b_, which represent the beginning and end DBH in each model year and are constrained by the first and last dendrometer band measurements. 
The model was fit in R v4.0 using the functions developed in the *Rdendrom* package[@mcmahon_general_2015]. 
These functions take the time-series of manual dendrometer band measurements and return maximum-likelihood optimized values of the above five parameters that best predict DBH for each day of year. 
An advantage of this approach is that short-term shrinkage and swelling associated with rain events [@mcmahon_general_2015;@herrmann_tree_2016] and measurement errors show up as residual variation and do not unduly influence the parameters of interest. 
Having fit the model for each band-year of data, we then modeled DBH using these optimal parameter values in our logistic growth model and extracted the intra-annual growth variables of interest (Fig. `r schematic_summary`).

>>>>

We fit a five-parameter logistic growth model[@mcmahon_general_2015] to dendrometer band data from each tree-year to define stem growth milestones and growth rates (Fig. `r schematic_summary`). 
In particular, we model the observed diameter at breast height (DBH) on a given day of the year (DOY; *i.e.*, julian days) as: 
$$ DBH = L + \frac{K - L} { 1 + 1/\theta \cdot exp[-r(DOY - DOY_{ip}) / \theta] ^ \theta}$$ 
Here, _L_ and _K_ are lower and upper asymptotes of the model. 
*DOY~ip~* is the day of year where the inflection point in growth rate occurs,
_r_ shapes the slope of the curve at the inflection point,
and $\theta$ is a tuning parameter controlling the slope of the curve toward the upper asymptote.
This allows an asymmetric fit to the data, where the onset of growth can be estimated independent of the cessation of growth. 
When $\theta = 1$, *g~max~* occurs on *DOY~ip~*.
The model outputs two additional variables, _a_ and _b_, which represent the beginning and end DBH in each model year and are constrained by the first and last dendrometer band measurements. 
The model was fit in R v4.0 using the functions developed in the *Rdendrom* package[@mcmahon_general_2015]. 
These functions take the time-series of manual dendrometer band measurements and return maximum-likelihood optimized values of the above five parameters that best predict DBH for each day of year. 
An advantage of this approach is that short-term shrinkage and swelling associated with rain events [@mcmahon_general_2015;@herrmann_tree_2016] and measurement errors show up as residual variation and do not unduly influence the parameters of interest. 
Having fit the model for each band-year of data, we then modeled DBH using these optimal parameter values in our logistic growth model and extracted the intra-annual growth variables of interest (Fig. `r schematic_summary`).
